Principles for Open Data Curation: A Case Study with the New York City 311 Service Request Data

### 2024-06-17 manuscript/manuscript.tex
removed backslash in figure file names in includegraphics
Clean up of spaces and indentations for consistent presentation

\hyphenpenalty=1000

@@

\section{Introduction} \label{sec:intro}

In the early 21st century, the open data movement began to take shape, driven by the 
fundamental belief that freely accessible data can transform 
both societies and governments. This movement champions the principles of 
transparency, innovation, and public engagement. 
A landmark in this journey was the launch of the United States'
\href{https://www.data.gov}{Data.gov} portal in 2009, a pioneering
platform in making government data widely accessible. Shortly after,
the European Union followed suit, unveiling its
\href{https://data.europa.eu/euodp}{Open Data Portal} in 2012, further
cementing the movement's global reach. Furthermore, the World Bank's Open
Data initiative, initiated in 2010, stands out as a comprehensive
repository for global development data, available at
\href{https://data.worldbank.org}{World Bank Open Data}. 
These initiatives represent significant strides in democratizing data, 
in breaking barriers that once kept valuable information 
on government performance in silos. Their collective impact 
is profound, extending beyond mere data sharing to 
fostering a culture of openness that benefits individuals, 
communities, governments, and economies worldwide 
\citep{barns2016mine, wang2016adoption}.

@@

A note on naming conventions: dataset field names are typically 
named using ``snake\_case'' where each word in the 
variable name is written in lowercase, and 
words are separated by underscores, e.g 
created\_date, complaint\_type, zip\_code, etc..

\section{NYC 311 System History} \label{sec:history}

The NYC 311 service, a critical component of New York City's public
engagement and service response framework, serves as a centralized hub
for non-emergency inquiries and requests. Introduced in 2003, the NYC
311 system was designed to streamline the city's response to
non-emergency issues, ranging from noise complaints to street
maintenance requests. The evolution of this system can be traced from
its initial implementation as a simple inquiry channel via telephone to a
comprehensive data management system that handles millions of requests
annually. Each inquiry is logged as a complaint in the 311 system.
Key milestones in the 311 system development include:

@@

Additional Open Data projects can be found 
at \href{https://opendata.cityofnewyork.us/projects/}{NYC Open Data Project Gallery}
and at \href{https://beta.nyc/beta/products/}{\textbeta etaNYC} products and tools.

Despite its success, the system constantly faces challenges such as:

@@

These and other challenges are continually addressed by the various Agency open data managers
as well as the 
\href{https://www.nyc.gov/content/oti/pages/}{NYC Office of Technology and Innovation (OTI)} 
which provides the technology support for the open data system.  
Many of the underlying datasets are updated daily, usually running 
1-2 days behind from creation. However, a number of datasets 
are updated monthly (such as restaurant inspections), and some are 
not updated for one or more years (such as community board 
boundaries). This poses problems when merging datasets with 
others, as would be expected. Typically, the data must be exported 
and manipulated in another, external program. The authors found 
this to be the case with the large 311 dataset employed.

@@

\section{10-year Analysis \& Trends} \label{sec:trends}

Since the NYC 311 System went live in 2003, the usage has grown at an annual
rate of 4.6\%. As part of this study, we also looked at a 10-year period (2014-2023) to
observe system performance over a longer time-frame . During that time-frame, 
the number of 311 SRs grew 50\%. The summer  and early fall months 
make up all of the Top 5 SR counts by day. 
The all-time daily high for SRs, 24,415 SRs (a full 9.8 \textsigma's from the 10-year mean), 
occurred on Tuesday, 2020-08-04.
The 10-year low count occurs on Christmas Day, 2016, a Sunday. 
Note also that the difference between 
the peak day (24,415) and the minimum day (2965) varies by 8X, creating a 
challenging system performance issue. This timeframe (2014-2023) 
also includes the COVID pandemic,  
which reached a peak in the July-September 2020 timeframe, 
accompanied by additional spikes in the 2\textsuperscript{nd}half of 2021 as well.

@@

\begin{table}[tbp]
    \centering
    \normalsize
    \caption{Top 5 and Bottom 5 Days for SRs}
    \begin{tabular}{@{}lll lll@{}}
        \toprule
        \multicolumn{3}{c}{Top 5 Days} & \multicolumn{3}{c}{Bottom 5 Days} \\
        \cmidrule(r){1-3} \cmidrule(l){4-6}
        Date & Count & Rank & Date & Count & Rank \\
        \midrule
        2020-08-04 & 24,415 & 1 & 2014-12-25 & 2965 & 1 \\
        2020-08-05 & 19,560 & 2 & 2015-12-25 & 3247 & 2 \\
        2023-09-29 & 17,962 & 3 & 2014-04-20 & 3287 & 3 \\
        2020-07-05 & 16,916 & 4 & 2015-12-26 & 3405 & 4 \\
        2020-06-21 & 15,883 & 5 & 2016-11-06 & 3419 & 5 \\
        \bottomrule
    \end{tabular}
    \label{tab:top-10-counts}
\end{table}

@@

\begin{table}[tbp]
    \centering
    \normalsize
    \caption{Noise-related complaints\_type(s) by count with Agency}
    \begin{tabular}{@{}lS[table-format=7.0,round-mode=places,
    round-precision=0]S[table-format=2.2,round-mode=places,
    round-precision=2]l@{}} % 'l' for left-aligned, 'S' for siunitx number alignment
        \toprule
        \textbf{complaint\_type} & \textbf{Count} & \textbf{Percentage} & \textbf{Agency} \\ 
        \midrule
        NOISE - RESIDENTIAL        & 675502 & 10.56 & NYPD  \\ 
        NOISE - STREET/SIDEWALK    & 300507 &  4.70 & NYPD  \\ 
        NOISE - COMMERCIAL         & 128892 &  2.02 & NYPD  \\ 
        NOISE - VEHICLE            & 115956 &  1.81 & NYPD  \\ 
        NOISE                      & 100413 &  1.57 & DEP   \\ 
        NOISE - HELICOPTER         &  85345 &  1.33 & EDC   \\ 
        NOISE - PARK               &  16620 &  0.26 & NYPD  \\ 
        NOISE - HOUSE OF WORSHIP   &   2757 &  0.04 & NYPD  \\ 
        \bottomrule
    \end{tabular}
    \label{tab:noisecomplaints}
\end{table}

@@

\section{Data Cleansing Issues} \label{sec:issues}

What is data cleansing?  Wikipedia 
\href{https://en.wikipedia.org/wiki/Data_cleansing}{Data Cleansing} offers a good definition.

``Data cleansing or data cleaning is the process of detecting and 
correcting (or removing) corrupt or inaccurate records from a record set, 
table, or database and refers to identifying incomplete, incorrect, 
inaccurate or irrelevant parts of the data 
and then replacing, modifying, or deleting the dirty or coarse data.''

@@

\section{Structural Issues}\label{sec:structural}

Structural issues in the context of data cleansing refers to issues 
related to the how data is organized, formatted, or structured within a dataset. Structural issues 
can make it difficult to analyze the data effectively. Some common 
structural concerns in this 311 SR 2022-2023 dataset include:

@@

\section{Validating data types}\label{sec:datatypes}
Fortunately, both the Data Dictionary and the portal column manager 
do a good job of indicating the data type.
As indicated by the red underline in Figure \ref{fig:computed-columns}, 
there is a small icon next to the field name which indicates the 
data type of the field. The stylized capital ``T'' indicates ``text'', the ``\#'' sign indicates 
``numeric'', a calendar icon indicates a ``date'' field, and the pin on a 
map symbol indicates a ``geospatial'' field.

The following fields were checked for their specified data types:

@@

\section{Blank \& N/A data}\label{sec:blanks}

Understanding the absence of data by field is an important factor 
when undertaking analysis. For example, if you wanted to see if the SRs were
closed before or after their due\_date, you would be challenged 
as 99.57\% of the due\_date field is blank. (Although with such a large dataset
that still leaves 24,411 values, all of which are ``N/A'. Which calls 
into question...are these missing values, or is the concept of a
due\_date simply not applicable for the vast majority of SRs?

@@

\begin{table}[tbp]
\centering
\normalsize
\caption{Blank and N/A entries by Field}
\resizebox{0.8\linewidth}{!}{%
\begin{tabular}{lrrrl}
	\toprule
	\textbf{Field} & \textbf{Total Empty} & \textbf{Percent Empty} 
	& \textbf{N/A Count} \\
	\midrule
		taxi\_company\_borough & 6391341 & 99.94 & 0 \\
		road\_ramp & 6379080 & 99.75 & 0 \\
		vehicle\_type & 6376266 & 99.71 & 0 \\
		due\_date & 6370497 & 99.62 & 6370497 \\
		bridge\_highway\_direction & 6367626 & 99.57 & 0 \\
		bridge\_highway\_name & 6340197 & 99.14 & 0 \\
		bridge\_highway\_segment & 6340190 & 99.14 & 0 \\
		taxi\_pick\_up\_location & 6329954 & 98.98 & 0 \\
		facility\_type & 5966588 & 93.30 & 0 \\
		landmark & 2635736 & 41.22 & 0 \\
		intersection\_street\_1 & 2143847 & 33.52 & 0 \\
		intersection\_street\_2 & 2140707 & 33.48 & 0 \\
		cross\_street\_1 & 1846594 & 28.88 & 0 \\
		cross\_street\_2 & 1846008 & 28.87 & 0 \\
		location\_type & 798775 & 12.49 & 0 \\
		bbl & 768109 & 12.01 & 0 \\
		city & 332086 & 5.19 & 0 \\
		street\_name & 280463 & 4.39 & 0 \\
		incident\_address & 280268 & 4.38 & 0 \\
		closed\_date & 248839 & 3.89 & 248839 \\
		resolution\_description & 132333 & 2.07 & 0 \\
		zip\_codes & 123167 & 1.93 & 0 \\
		borough\_boundaries & 101453 & 1.59 & 0 \\
		community\_districts & 101445 & 1.59 & 0 \\
		city\_council\_districts & 101445 & 1.59 & 0 \\
		police\_precincts & 101445 & 1.59 & 0 \\
		police\_precinct & 101416 & 1.59 & 0 \\
		latitude & 99778 & 1.56 & 0 \\
		longitude & 99778 & 1.56 & 0 \\
		location & 99778 & 1.56 & 0 \\
		x\_coordinate\_state\_plane & 99668 & 1.56 & 0 \\
		y\_coordinate\_state\_plane & 98822 & 1.55 & 0 \\
		resolution\_action\_updated\_date & 91399 & 1.43 & 91399 \\
		incident\_zip & 83220 & 1.30 & 2 \\
		descriptor & 74645 & 1.17 & 0 \\
		address\_type & 38655 & 0.60 & 0 \\
		unique\_key & 0 & 0.00 & 0 \\
		agency & 0 & 0.00 & 0 \\
		agency\_name & 0 & 0.00 & 0 \\
		complaint\_type & 0 & 0.00 & 0 \\
		status & 0 & 0.00 & 0 \\
		community\_board & 0 & 0.00 & 0 \\
		borough & 0 & 0.00 & 0 \\
		open\_data\_channel\_type & 0 & 0.00 & 0 \\
		park\_facility\_name & 0 & 0.00 & 0 \\
		park\_borough & 0 & 0.00 & 0 \\
		created\_date & 0 & 0.00 & 0 \\
	\bottomrule
\end{tabular}
}
\end{table}

%\FloatBarrier % Ensure all floats before this point are processed before moving on

@@

		\subsubsection{Case Study: Noise Complaints by Zip Code}
		{label{sec:case-study-zip-codes}
		\textbf{Scenario:} The NYC Office of Nightlife (ONL) wants to 
		know ``What are the top 10 zip codes for Noise Complaints (all 8 types) 
		over the last two years?'' The goal is to assess the impact of the 
		recent NYC effort looking to promote a safe and vibrant 
		nightlife scene in NYC while seeking to ease strained 
		relations between bar and club owners.

@@

\begin{table}[tbp]
    \centering
    \normalsize
    \caption{Comparison of Top Ten Zip Codes Lists}
    \begin{tabular}{@{}lS[table-format=6.0,round-mode=places,
    round-precision=0]c@{\hskip 0.5cm}@{}lS[table-format=6.0,
    round-mode=places,round-precision=0]c@{}}
        \toprule
        \multicolumn{3}{c}{\textbf{zip\_codes}} & \multicolumn{3}{c}{\textbf{incident\_zip}} \\
        \cmidrule(r){1-3} \cmidrule(l){4-6}
        \textbf{Zip Code} & \textbf{Count} & \textbf{Valid?} 
        & \textbf{Zip Code} & \textbf{Count} & \textbf{Valid?} \\
        \midrule
        11275 & 104556 & FALSE & 10466 & 104562 & TRUE \\
        12420 & 27503 & TRUE & 10023 & 27972 & TRUE \\
        12428 & 26564 & TRUE & 10031 & 25548 & TRUE \\
        10935 & 25508 & FALSE & 10457 & 25066 & TRUE \\
        10934 & 23448 & FALSE & 10453 & 24752 & TRUE \\
        10931 & 22381 & TRUE & 10456 & 24751 & TRUE \\
        10930 & 22121 & TRUE & 10452 & 22527 & TRUE \\
        17613 & 21963 & FALSE & 10025 & 21705 & TRUE \\
        10936 & 21707 & FALSE & 10458 & 21689 & TRUE \\
        11606 & 21435 & FALSE & 10032 & 20622 & TRUE \\
        \bottomrule
    \end{tabular}
    \label{tab:zipcodes}
\end{table}

@@

	\begin{table}[tbp]
	\normalsize
	\centering
	\caption{Comparison of the fields police\_precinct and police\_precincts}
		\begin{tabular}{ccccc}
		\toprule
		\textbf{Invalid Precinct} & \textbf{police\_precinct count} 
		& \textbf{police\_precincts count} & \textbf{Delta}\\
		\midrule
			62 & 151808 & 151720 & 88 \\
			72 & 148881 & 148796 & 85 \\
			67 & 133171 & 133115 & 56 \\
			64 & 111236 & 111221 & 15 \\
			60 & 105252 & 105317 & -65 \\
			65 & 103334 & 103306 & 28 \\
			53 & 100097 & 100074 & 23 \\
			73 & 97746 & 97755 & -9 \\
			68 & 94204 & 94254 & -50 \\
			66 & 93654 & 93848 & -194 \\
	\bottomrule
	\end{tabular}
	\label{tab:comparison-precincts-diff}
	\end{table}

	A graphic representation of the invalid police\_precincts plotted 
	by Agency show a near perfect match to the distribution of the 
	entire 6.4 million SRs as seen in Figure \ref{fig:SR_counts_by_Agency}. 
	Note that the ``big six'' Agencies account for 90\% of these invalid 
	precinct errors, identical to the overall distribution of all SRs. 
	(The same is true for a similar graph of by the police\_precint field.) 
	This strongly suggests that this is a systemic problem, 
	almost certainly caused by the computation method applied for these two fields.

@@

	\subsection{Issues with Community Boards}
	Community Boards are an important aspect of NYC government. They are the 
	most local, grassroots form of City government, and a vital connection 
	between communities and elected officials and City agencies. 
	They are organized by the five Boroughs of NYC and represented	
	in this dataset as ``\#\#-Borough'', e.g. ``10 Bronx'' and report to the 
	various Borough Presidents in NYC. As such, they play an important 
	role in the quality of life for all New Yorkers. There are 59 community 
	boards throughout the City: 12 Community Districts in Bronx, 18 in Brooklyn, 
	12 in Manhattan, 14 in Queens and 3 in Staten Island. The Community 
	Boards are used frequently as a way to measure distribution of 
	City services throughout the five boroughs, so correctness is important. 
	NYC Borough Presidents in particular are focused on the services 
	offered at the Community Board level within their respective Borough.

@@

		\begin{table}[tbp]
		\centering
		\normalsize
		\caption{Top Ten Invalid 'community\_board' Values}
		\begin{tabular}{rlr}
		\toprule
		\textbf{Rank} & \textbf{Invalid CB} & \textbf{Count} \\
			\midrule
				1 & 64 MANHATTAN & 7144 \\
				2 & 83 QUEENS & 5133 \\
				3 & 55 BROOKLYN & 3327 \\
				4 & 81 QUEENS & 3180 \\
				5 & 80 QUEENS & 2407 \\
				6 & 26 BRONX & 2364 \\
				7 & 28 BRONX & 1128 \\
				8 & 82 QUEENS & 1054 \\
				9 & 95 STATEN ISLAND & 515 \\
				10 & 27 BRONX & 514 \\
			\bottomrule
		\end{tabular}
		\end{table}

@@

	\subsection{Issues with Community Districts}
	Another of the computed fields is community\_districts. Community Districts 
	are the boundaries for the Community Boards, but unlike how 
	Community Boards are an arm of the government of New York City, 
	the Community District is used by the Department of City Planning 
	(DCP) for purposes of environmental, socio-economic, and demographic 
	purposes. DCP is the NYC''s primary land use agency and is 
	instrumental in the City's physical framework.  It is a geographical 
	division rather than a local government division. And like Community 
	Boards, the Community Districts are frequently cited as a means to 
	track zoning, housing, community facilities, and waterfront/open/public spaces. 
	It is another of the NYC well-tracked measures.

@@

\subsection{created\_date and closed\_date(s) -- Negative Duration}
Let's look first at the created\_date and closed\_date fields. You might 
expect that these two fields are automatically populated by the 
SR application software, either the software used at the 311 organization 
or that used at the Agencies that are integrated with the 311 system; 
and typically that is indeed the case. In such a case, the act of creating an 
SR would automatically trigger an entry in created\_date based on a 
system reference clock. Thus the created\_date is associated with 
setting an SR status to ``new'' or ``open''. And similarly when an SR's 
status is changed to ``closed'' then the closed\_date would be populated.

@@

\subsubsection{Case Study: Homeless Person Assistance}
		\textbf{Scenario:} The Dept. of Homeless Services wants to 
		know ``How quickly are 311 calls for ``Homeless Person Assistance 
		over the past two years?'' resolved. This is a typical request made by 
		both the public and City government. It's a performance metric along 
		the lines of ``How quickly is my Agency responding to ``critical'' 
		requests, and does that performance vary by Borough, Zip Code, etc. 
		As an analyst, this appears to be a routine effort:

@@

\subsection{latitude \& longitude and the location fields}  The location field is a 
pure concatenation of the latitude and longitude fields, with a comma and 
parenthesis added. Example:

>>>>

\hyphenpenalty=750

@@

\section{Introduction} \label{sec:intro}
In the early 21st century, the open data movement began to take shape, driven by the 
fundamental belief that freely accessible data can transform 
both societies and governments. This movement champions the principles of 
transparency, innovation, and public engagement. 
A landmark in this journey was the launch of the United States'
\href{https://www.data.gov}{Data.gov} portal in 2009, a pioneering
platform in making government data widely accessible. Shortly after,
the European Union followed suit, unveiling its
\href{https://data.europa.eu/euodp}{Open Data Portal} in 2012, further
cementing the movement's global reach. Furthermore, the World Bank's Open
Data initiative, initiated in 2010, stands out as a comprehensive
repository for global development data, available at
\href{https://data.worldbank.org}{World Bank Open Data}. 
These initiatives represent significant strides in democratizing data, 
in breaking barriers that once kept valuable information 
on government performance in silos. Their collective impact 
is profound, extending beyond mere data sharing to 
fostering a culture of openness that benefits individuals, 
communities, governments, and economies worldwide 
\citep{barns2016mine, wang2016adoption}.

@@

A note on naming conventions: dataset field names are typically 
named using ``snake\_case'' where each word in the variable name 
is written in lowercase, and words are separated by 
underscores, e.g. created\_date, complaint\_type, zip\_code, etc..

\section{NYC 311 System History} \label{sec:history}
The NYC 311 service, a critical component of New York City's public
engagement and service response framework, serves as a centralized hub
for non-emergency inquiries and requests. Introduced in 2003, the NYC
311 system was designed to streamline the city's response to
non-emergency issues, ranging from noise complaints to street
maintenance requests. The evolution of this system can be traced from
its initial implementation as a simple inquiry channel via telephone to a
comprehensive data management system that handles millions of requests
annually. Each inquiry is logged as a complaint in the 311 system.
Key milestones in the 311 system development include:

@@

Additional Open Data projects can be found 
at \href{https://opendata.cityofnewyork.us/projects/}{NYC Open Data Project Gallery}
and at \href{https://beta.nyc/beta/products/}{\textbeta etaNYC} products and tools. 
Despite its success, the system constantly faces challenges such as:

@@

These and other challenges are continually addressed by the various 
Agency open data managers as well as the 
\href{https://www.nyc.gov/content/oti/pages/}
{NYC Office of Technology and Innovation (OTI)} which provides the 
technology support for the open data system. Many of the underlying 
datasets are updated daily, usually running 1-2 days behind 
from creation. However, a number of datasets are updated 
monthly (such as restaurant inspections), and some are not 
updated for one or more years (such as community board 
boundaries). This poses problems when merging datasets with 
others, as would be expected. Typically, the data must be exported 
and manipulated in another, external program. The authors found 
this to be the case with the large 311 dataset employed.

@@

\section{10-year Analysis \& Trends} \label{sec:trends}
Since the NYC 311 System went live in 2003, the usage has grown at an annual
rate of 4.6\%. As part of this study, we also looked at a 10-year period (2014-2023) to
observe system performance over a longer time-frame . During that time-frame, 
the number of 311 SRs grew 50\%. The summer  and early fall months 
make up all of the Top 5 SR counts by day. The all-time daily 
high for SRs, 24,415 SRs (a full 9.8 \textsigma's from the 10-year mean), 
occurred on Tuesday, 2020-08-04. The 10-year low count occurs 
on Christmas Day, 2016, a Sunday. Note also that the difference between 
the peak day (24,415) and the minimum day (2965) varies by 8X, creating a 
challenging system performance issue. This timeframe (2014-2023) 
also includes the COVID pandemic, which reached a peak in 
the July-September 2020 timeframe, accompanied by additional 
spikes in the 2\textsuperscript{nd}half of 2021 as well.

@@

\begin{table}[tbp]
    \centering
    \caption{Top 5 and Bottom 5 Days for SRs}
	    \begin{tabular}{@{}lll lll@{}}
	        \toprule
	        \multicolumn{3}{c}{Top 5 Days} & \multicolumn{3}{c}{Bottom 5 Days} \\
	        \cmidrule(r){1-3} \cmidrule(l){4-6}
	        Date & Count & Rank & Date & Count & Rank \\
	        \midrule
	        2020-08-04 & 24,415 & 1 & 2014-12-25 & 2965 & 1 \\
	        2020-08-05 & 19,560 & 2 & 2015-12-25 & 3247 & 2 \\
	        2023-09-29 & 17,962 & 3 & 2014-04-20 & 3287 & 3 \\
	        2020-07-05 & 16,916 & 4 & 2015-12-26 & 3405 & 4 \\
	        2020-06-21 & 15,883 & 5 & 2016-11-06 & 3419 & 5 \\
	        \bottomrule
	    \end{tabular}
    \label{tab:top-10-counts}
\end{table}

@@

\begin{table}[tbp]
    \centering
    \caption{Noise-related complaints\_type(s) by count with Agency}
	    \begin{tabular}{@{}lS[table-format=7.0,round-mode=places,
	    round-precision=0]S[table-format=2.2,round-mode=places,
	    round-precision=2]l@{}} % 'l' for left-aligned, 'S' for siunitx number alignment
	        \toprule
	        \textbf{complaint\_type} & \textbf{Count} & \textbf{Percentage} & \textbf{Agency} \\ 
	        \midrule
	        Noise - Residential        & 675502 & 10.56 & NYPD  \\ 
	        Noise - Street/Sidewalk    & 300507 &  4.70 & NYPD  \\ 
	        Noise - Commercial         & 128892 &  2.02 & NYPD  \\ 
	        Noise - Vehicle            & 115956 &  1.81 & NYPD  \\ 
	        Noise                      & 100413 &  1.57 & DEP   \\ 
	        Noise - Helicopter         &  85345 &  1.33 & EDC   \\ 
	        Noise - Park               &  16620 &  0.26 & NYPD  \\ 
	        Noise - House of Worship &   2757 &  0.04 & NYPD  \\ 
	        \bottomrule
	    \end{tabular}
    \label{tab:noisecomplaints}
\end{table}

@@

\section{Data Cleansing Issues} \label{sec:issues}
What is data cleansing?  Wikipedia 
\href{https://en.wikipedia.org/wiki/Data_cleansing}{Data Cleansing} offers a good definition.

\begin{quote}\textit{Data cleansing or data cleaning is the process of detecting and 
correcting (or removing) corrupt or inaccurate records from a record set, 
table, or database and refers to identifying incomplete, incorrect, 
inaccurate or irrelevant parts of the data and then replacing, 
modifying, or deleting the dirty or coarse data.}
\end{quote}

@@

\section{Structural Issues}\label{sec:structural}
Structural issues in the context of data cleansing refers to issues 
related to the how data is organized, formatted, or structured within a dataset. Structural issues 
can make it difficult to analyze the data effectively. Some common 
structural concerns in this 311 SR 2022-2023 dataset include:

@@

\section{Validating data types}\label{sec:datatypes}
Fortunately, both the Data Dictionary and the portal column manager 
do a good job of indicating the data type. As indicated by 
the red underline in Figure \ref{fig:computed-columns}, there 
is a small icon next to the field name which indicates the 
data type of the field. The stylized capital ``T'' indicates ``text'', the ``\#'' sign indicates 
``numeric'', a calendar icon indicates a ``date'' field, and the pin on a 
map symbol indicates a ``geospatial'' field. The following fields were 
checked for their specified data types:

@@

\section{Blank \& N/A data}\label{sec:blanks}
Understanding the absence of data by field is an important factor 
when undertaking analysis. For example, if you wanted to see if the SRs were
closed before or after their due\_date, you would be challenged 
as 99.57\% of the due\_date field is blank. (Although with such a large dataset
that still leaves 24,411 values, all of which are ``N/A'. Which calls 
into question...are these missing values, or is the concept of a
due\_date simply not applicable for the vast majority of SRs?

@@

\begin{table}[tbp]
\centering
\caption{Blank and N/A entries by Field}
\resizebox{0.8\linewidth}{!}{%
	\begin{tabular}{lrrrl}
		\toprule
		\textbf{Field} & \textbf{Total Empty} & \textbf{Percent Empty} 
		& \textbf{N/A Count} \\
		\midrule
			taxi\_company\_borough & 6391341 & 99.94 & 0 \\
			road\_ramp & 6379080 & 99.75 & 0 \\
			vehicle\_type & 6376266 & 99.71 & 0 \\
			due\_date & 6370497 & 99.62 & 6370497 \\
			bridge\_highway\_direction & 6367626 & 99.57 & 0 \\
			bridge\_highway\_name & 6340197 & 99.14 & 0 \\
			bridge\_highway\_segment & 6340190 & 99.14 & 0 \\
			taxi\_pick\_up\_location & 6329954 & 98.98 & 0 \\
			facility\_type & 5966588 & 93.30 & 0 \\
			landmark & 2635736 & 41.22 & 0 \\
			intersection\_street\_1 & 2143847 & 33.52 & 0 \\
			intersection\_street\_2 & 2140707 & 33.48 & 0 \\
			cross\_street\_1 & 1846594 & 28.88 & 0 \\
			cross\_street\_2 & 1846008 & 28.87 & 0 \\
			location\_type & 798775 & 12.49 & 0 \\
			bbl & 768109 & 12.01 & 0 \\
			city & 332086 & 5.19 & 0 \\
			street\_name & 280463 & 4.39 & 0 \\
			incident\_address & 280268 & 4.38 & 0 \\
			closed\_date & 248839 & 3.89 & 248839 \\
			resolution\_description & 132333 & 2.07 & 0 \\
			zip\_codes & 123167 & 1.93 & 0 \\
			borough\_boundaries & 101453 & 1.59 & 0 \\
			community\_districts & 101445 & 1.59 & 0 \\
			city\_council\_districts & 101445 & 1.59 & 0 \\
			police\_precincts & 101445 & 1.59 & 0 \\
			police\_precinct & 101416 & 1.59 & 0 \\
			latitude & 99778 & 1.56 & 0 \\
			longitude & 99778 & 1.56 & 0 \\
			location & 99778 & 1.56 & 0 \\
			x\_coordinate\_state\_plane & 99668 & 1.56 & 0 \\
			y\_coordinate\_state\_plane & 98822 & 1.55 & 0 \\
			resolution\_action\_updated\_date & 91399 & 1.43 & 91399 \\
			incident\_zip & 83220 & 1.30 & 2 \\
			descriptor & 74645 & 1.17 & 0 \\
			address\_type & 38655 & 0.60 & 0 \\
			unique\_key & 0 & 0.00 & 0 \\
			agency & 0 & 0.00 & 0 \\
			agency\_name & 0 & 0.00 & 0 \\
			complaint\_type & 0 & 0.00 & 0 \\
			status & 0 & 0.00 & 0 \\
			community\_board & 0 & 0.00 & 0 \\
			borough & 0 & 0.00 & 0 \\
			open\_data\_channel\_type & 0 & 0.00 & 0 \\
			park\_facility\_name & 0 & 0.00 & 0 \\
			park\_borough & 0 & 0.00 & 0 \\
			created\_date & 0 & 0.00 & 0 \\
		\bottomrule
	\end{tabular}
}
\end{table}

@@

\subsubsection{Case Study: Noise Complaints by Zip Code}
\label{sec:case-study-zip-codes}
		\textbf{Scenario:} The NYC Office of Nightlife (ONL) wants to 
		know ``What are the top 10 zip codes for Noise Complaints (all 8 types) 
		over the last two years?'' The goal is to assess the impact of the 
		recent NYC effort looking to promote a safe and vibrant 
		nightlife scene in NYC while seeking to ease strained 
		relations between bar and club owners.

@@

\begin{table}[tbp]
    \centering
    \caption{Comparison of Top Ten Zip Codes Lists}
	    \begin{tabular}{@{}lS[table-format=6.0,round-mode=places,
	    round-precision=0]c@{\hskip 0.5cm}@{}lS[table-format=6.0,
	    round-mode=places,round-precision=0]c@{}}
		\toprule
	 	\multicolumn{3}{c}{\textbf{zip\_codes}} & \multicolumn{3}{c}{\textbf{incident\_zip}} \\
	      \cmidrule(r){1-3} \cmidrule(l){4-6}
	      \textbf{Zip Code} & \textbf{Count} & \textbf{Valid?} 
	      & \textbf{Zip Code} & \textbf{Count} & \textbf{Valid?} \\
	      \midrule
	        11275 & 104556 & FALSE & 10466 & 104562 & TRUE \\
	        12420 & 27503 & TRUE & 10023 & 27972 & TRUE \\
	        12428 & 26564 & TRUE & 10031 & 25548 & TRUE \\
	        10935 & 25508 & FALSE & 10457 & 25066 & TRUE \\
	        10934 & 23448 & FALSE & 10453 & 24752 & TRUE \\
	        10931 & 22381 & TRUE & 10456 & 24751 & TRUE \\
	        10930 & 22121 & TRUE & 10452 & 22527 & TRUE \\
	        17613 & 21963 & FALSE & 10025 & 21705 & TRUE \\
	        10936 & 21707 & FALSE & 10458 & 21689 & TRUE \\
	        11606 & 21435 & FALSE & 10032 & 20622 & TRUE \\
	      \bottomrule
	    	\end{tabular}
 	\label{tab:zipcodes}
\end{table}

@@

	\begin{table}[tbp]
	\centering
	\caption{Comparison of the fields police\_precinct and police\_precincts}
		\begin{tabular}{ccccc}
		\toprule
		\textbf{Invalid Precinct} & \textbf{police\_precinct count} 
		& \textbf{police\_precincts count} & \textbf{Delta}\\
		\midrule
				62 & 151808 & 151720 & 88 \\
				72 & 148881 & 148796 & 85 \\
				67 & 133171 & 133115 & 56 \\
				64 & 111236 & 111221 & 15 \\
				60 & 105252 & 105317 & -65 \\
				65 & 103334 & 103306 & 28 \\
				53 & 100097 & 100074 & 23 \\
				73 & 97746 & 97755 & -9 \\
				68 & 94204 & 94254 & -50 \\
				66 & 93654 & 93848 & -194 \\
		\bottomrule
		\end{tabular}
	\label{tab:comparison-precincts-diff}
	\end{table}

	A graphic representation of the invalid police\_precincts plotted by 
	Agency show a near perfect match to the distribution of the entire 
	6.4 million SRs as seen in Figure \ref{fig:SR_counts_by_Agency}. Note 
	that the ``big six'' Agencies account for 90\% of these invalid precinct 
	errors, identical to the overall distribution of all SRs. (The same is true 
	for a similar graph of by the police\_precint field.) This strongly 
	suggests that this is a systemic problem, most likely caused by the 
	computation method applied for these two fields.

@@

\subsection{Issues with Community Boards}
\label{sec:communityboards}
	Community Boards are an important aspect of NYC government. They are the 
	most local, grassroots form of City government, and a vital connection 
	between communities and elected officials and City agencies. 
	They are organized by the five Boroughs of NYC and represented	
	in this dataset as ``\#\#-Borough'', e.g. ``10 Bronx'' and report to the 
	various Borough Presidents in NYC. As such, they play an important 
	role in the quality of life for all New Yorkers. There are 59 community 
	boards throughout the City: 12 Community Districts in Bronx, 18 in Brooklyn, 
	12 in Manhattan, 14 in Queens and 3 in Staten Island. The Community 
	Boards are used frequently as a way to measure distribution of 
	City services throughout the five boroughs, so correctness is important. 
	NYC Borough Presidents in particular are focused on the services 
	offered at the Community Board level within their respective Borough.

@@

		\begin{table}[tbp]
		\centering
		\caption{Top Ten Invalid 'community\_board' Values}
			\begin{tabular}{rlr}
			\toprule
			\textbf{Rank} & \textbf{Invalid CB} & \textbf{Count} \\
				\midrule
					1 & 64 MANHATTAN & 7144 \\
					2 & 83 QUEENS & 5133 \\
					3 & 55 BROOKLYN & 3327 \\
					4 & 81 QUEENS & 3180 \\
					5 & 80 QUEENS & 2407 \\
					6 & 26 BRONX & 2364 \\
					7 & 28 BRONX & 1128 \\
					8 & 82 QUEENS & 1054 \\
					9 & 95 STATEN ISLAND & 515 \\
					10 & 27 BRONX & 514 \\
				\bottomrule
			\end{tabular}
		\end{table}

@@

\subsection{Issues with Community Districts}
\label{sec:communitydistrict}
	Another of the computed fields is community\_districts. Community Districts 
	are the boundaries for the Community Boards, but unlike how 
	Community Boards are an arm of the government of New York City, 
	the Community District is used by the Department of City Planning 
	(DCP) for purposes of environmental, socio-economic, and demographic 
	purposes. DCP is the NYC''s primary land use agency and is 
	instrumental in the City's physical framework.  It is a geographical 
	division rather than a local government division. And like Community 
	Boards, the Community Districts are frequently cited as a means to 
	track zoning, housing, community facilities, and waterfront/open/public spaces. 
	It is another of the NYC well-tracked measures.

@@

\subsection{created\_date and closed\_date(s) -- Negative Duration}
\label{sec:negativeduration}
	Let's look first at the created\_date and closed\_date fields. You might 
	expect that these two fields are automatically populated by the 
	SR application software, either the software used at the 311 organization 
	or that used at the Agencies that are integrated with the 311 system; 
	and typically that is indeed the case. In such a case, the act of creating an 
	SR would automatically trigger an entry in created\_date based on a 
	system reference clock. Thus the created\_date is associated with 
	setting an SR status to ``new'' or ``open''. And similarly when an SR's 
	status is changed to ``closed'' then the closed\_date would be populated.

@@

\subsubsection{Case Study: Homeless Person Assistance}
\label{sec:homlessassistance}
	\textbf{Scenario:} The Dept. of Homeless Services wants to 
	know ``How quickly are 311 calls for ``Homeless Person Assistance 
	over the past two years?'' resolved. This is a typical request made by 
	both the public and City government. It's a performance metric along 
	the lines of ``How quickly is my Agency responding to ``critical'' 
	requests, and does that performance vary by Borough, Zip Code, etc. 
	As an analyst, this appears to be a routine effort:

@@

\subsection{latitude \& longitude and the location fields}
\label{sec:latlong}
The location field is a pure concatenation of the latitude 
and longitude fields, with a comma and parenthesis 
added. Example:
